---
title: "04. Get Labware metadata"
author: "DHJ"
date: "2 7 2019"
output: html_document
---

## Setup

### 1a. Libraries
```{r}
library(dplyr)
library(niRvana)
library(ggplot2)
library(tidyr)
```

### 1b. Connect to database
```{r}
connect_niva()
```

### 1c. Functions
```{r}

# From selected records in LABWARE_CHECK_SAMPLE, check data in Merete's dataset
get_matrix_species <- function(df_nivabasen){
  limscodes <- df_nivabasen %>% 
    pull(TEXT_ID)
  dat %>%
    filter(LIMS %in% limscodes) %>%
    count(Matrix_orig, Species)
}

# From selected records in LABWARE_CHECK_SAMPLE, change LIMS code to 12 characters and check data
#   in Merete's dataset
get_matrix_species_12char <- function(df_nivabasen){
  limscodes <- df_nivabasen %>% 
    mutate(TEXT_ID_12 = paste0(substr(TEXT_ID, 1, 8), substr(TEXT_ID, 10, 13))) %>%
    pull(TEXT_ID_12)
  dat %>%
    filter(LIMS %in% limscodes) %>%
    count(Matrix_orig, Species)
}
```


## 2. Data
### Data from Merete (stacked data)
```{r}
dat <- readxl::read_excel("Input_data/001 PFAS stacked data table.xlsx", guess_max = 11000)
# A few errors in the 'PFASlength' variable
tab <- table(dat$LIMS)
length(tab)  # 2809
samplecodes <- names(tab)
```

### Check year column  
Should fit with LIMS column
```{r}
check <- as.numeric(substr(dat$LIMS,4,7)) == dat$Year
mean(check)  # should be 1
```

## 3a. Preamble  
Lots of the 2017-18 data lack species - can we get this from LIMS?
```{r}
table(substr(dat$LIMS,4,7), !is.na(dat$Species))
```

Lots of the 2017-18 data lack species - can we get this from LIMS?
```{r}
table(substr(dat$LIMS,4,7), !is.na(dat$Organ))
```


## 3b. Add columns from LABWARE_CHECK_SAMPLE to the Excel data   
### No of characters in stacked data
Almost all 2018 data lacking Species and Organ has 13 characters
```{r}
sel_year <- substr(dat$LIMS,4,7) %in% c("2015","2016","2017","2018")
cat("Has species:\n")
table(nchar(dat$LIMS[sel_year]), !is.na(dat$Species[sel_year]))   # Al
cat("\nHas organ:\n")
table(nchar(dat$LIMS[sel_year]), !is.na(dat$Organ[sel_year]))   # Al
```


### Get 12 and 13 character codes
```{r}
cat("13 character codes\n")
sel_13 <- nchar(dat$LIMS) == 13
tab <- table(dat$LIMS[sel_year & sel_13])
length(tab)
codes_13 <- names(tab)
head(codes_13)

cat("\n12 character codes\n")
sel_12 <- nchar(dat$LIMS) == 12
tab <- table(dat$LIMS[sel_year & sel_12])
length(tab)
codes_12_orig <- names(tab)
# codes_12_orig  # 27 numbers from 2902 and up
cat("- Originals: \n")
head(codes_12_orig)
```

### Make modified LIMS for join
```{r}
dat <- dat %>% 
  mutate(LIMS_modified = case_when(
    Year >= 2015 & nchar(LIMS) == 13 ~ LIMS,
    Year >= 2015 & nchar(LIMS) == 12 ~ paste0(substr(LIMS, 1, 8), "0", substr(LIMS, 9, 12)),
    TRUE ~ as.character(NA)    
  ))
```



### Get Labware data
```{r}
df_labware <- get_nivabase_selection(
    "TEXT_ID,PROSJEKT,SAMPLE_TYPE,SAMPLED_DATE,DESCRIPTION,AQUAMONITOR_ID,AQUAMONITOR_CODE,AQUAMONITOR_NAME,SPECIES,TISSUE,BIOTA_SAMPLENO",
    "LABWARE_CHECK_SAMPLE",
    "TEXT_ID",
    dat %>% filter(!is.na(LIMS_modified)) %>% pull(LIMS_modified) %>% unique(),
    values_are_text = TRUE)
nrow(df_labware)  # 1067

xtabs(~PROSJEKT, df_labware)
xtabs(~DESCRIPTION, df_labware)
```

```{r}
# SLOW:
# df <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where TEXT_ID like 'NR-2015%'")
# Faster:
df <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where TEXT_ID like 'NR-2015%' and PROSJEKT like '%MILKYS%'")
prosjekt <- unique(df$PROSJEKT)
xtabs(~PROSJEKT, df)

```



### Join data
```{r}
nrow(dat)
dat <- dat %>% 
  left_join(df_labware, by = c("LIMS_modified" = "TEXT_ID"))
nrow(dat)
```

## 4. Check result  
### Matrices
```{r}
dat %>%
  count(Matrix)
```


### Table by species
```{r}
dat %>%
  filter(Matrix %in% "biota" & Year >= 2015) %>%
  count(nchar(LIMS), Species, SPECIES)
```

### Table by tissue
```{r}
dat %>%
  filter(Matrix %in% "biota" & Year >= 2015) %>%
  count(nchar(LIMS), Organ, TISSUE)
```

### Species of those with unknown tissue
```{r}
dat %>%
  filter(Matrix %in% "biota" & Year >= 2015) %>%
  filter(is.na(Organ) & is.na(TISSUE)) %>%
  count(nchar(LIMS), Species, SPECIES)
```

## 5, Check codes of cod with unknown tissue  
```{r}
dat_sel <- dat %>%
  filter(Matrix %in% "biota" & Year >= 2015) %>%
  filter(is.na(Organ) & is.na(TISSUE)) %>%
  filter(Species %in% "Gadus morhua")

dat_sel %>%
  count(Project, Year, Description)
```


### Check numbers
```{r}
xtabs(~Year, dat_sel)

sampleno <- df_sel %>% pull(LIMS) %>% unique()
length(sampleno)
head(sampleno)
x <- as.numeric(substr(sampleno,9,14))
plot(x)
plot(x[x < 232000])
x[x < 230300]

```


## 6. Check one station/year with *unknown* tissue  
I.e., the LIMS number in Excel doesn't fit with TEXT_ID in LABWARE_CHECK_SAMPLE  
But we will also check in the BIOTA_ tables (turns out to find the LIMS number there!)  

### In Excel file  
```{r}
dat_example <- dat_sel %>%
  filter(Project %in% "MilKys" & Year %in% 2015 & Description %in% "13B")
dat_example
```

### Check same year/station in NIVADATABASE.LABWARE_CHECK_SAMPLE
```{r}
labware_example_yr <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where TEXT_ID like 'NR-2015%' and PROSJEKT like '%MILKYS%'")
xtabs(~PROSJEKT, labware_example_yr)
prosjekt <- unique(df$PROSJEKT)
# prosjekt
labware_example <- labware_example_yr %>% filter(grepl("13B", DESCRIPTION, ignore.case = TRUE))
labware_example
```


### Check same station in BIOTA_ tables   
#### First, get BIOTA data from Nivadatabasen
```{r}
df_projects <- get_projects()   # we call it 'df_projects' (the default name used by 'get_stations_from_project')
# Get a list of the stations in the CEMP_Biota project
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)
# Get all specimens collected at these stations (20 seconds or so)
df_specimens <- get_specimens_from_stationdata(df_stations)
df_2015 <- get_biota_chemistry(
  years = 2015, 
  specimendata = df_specimens, 
  stationdata = df_stations,
  report_samples = TRUE)
```


#### Check data  
AHA - LIMS in Excel file fits SAMPLE_ID in BIOTA_SAMPLES!    
```{r}
biota_example <- df_2015 %>% filter(STATION_CODE %in% "13B" & grepl("PFOS", NAME))
biota_example

```

#### Compare sample numbers    
```{r}
cat("From Excel file (LIMS):\n")
unique(dat_example$LIMS)
cat("\nFrom biota_samples (SAMPLE_ID):\n")
unique(biota_example$SAMPLE_ID)
```


## 7. Check one station/year with *known* tissue  
These are no problem, really, so this is just 'for the record'  

### In Excel file  
```{r}
dat_example <- dat %>%
  filter(Year %in% 2015 & SPECIES %in% "Perca fluviatilis")

dat_example %>%
  count(LIMS, Species, DESCRIPTION) 
```

### Check same year/station in NIVADATABASE.LABWARE_CHECK_SAMPLE
```{r}
labware_example <- get_nivabase_selection("*", "LABWARE_CHECK_SAMPLE", "TEXT_ID", unique(dat_example$LIMS),
                                          values_are_text = TRUE)
labware_example
```


### Check same station in BIOTA_ tables     

#### First, get BIOTA data from Nivadatabasen  
We start with the O-number (project number)  
```{r}
# get_nivabase_data("select * from NIVADATABASE.PROJECTS_O_NUMBERS where rownum < 10")
df_onum <- get_nivabase_data("select * from NIVADATABASE.PROJECTS_O_NUMBERS where O_NUMBER = '13223'")  # n = 1
df_projects <- get_nivabase_selection("*", "PROJECTS", "PROJECT_ID", df_onum$PROJECT_ID)                # n = 1
df_projects_stations <- get_nivabase_selection("*", "PROJECTS_STATIONS", "PROJECT_ID", df_projects$PROJECT_ID)  # n = 1
df_specimens <- get_specimens_from_stationdata(df_projects_stations)  # n = 255 
df_conc <- get_biota_chemistry(
  years = 2015, 
  specimendata = df_specimens, 
  stationdata = df_projects_stations,
  report_samples = TRUE)
```


#### Check the same subset of the data  
```{r}
biota_example <- df_conc %>% filter(LATIN_NAME %in% "Perca fluviatilis" & grepl("PFOS", NAME))
biota_example
```

#### Compare sample numbers
```{r}
cat("From Excel file (LIMS):\n")
unique(dat_example$LIMS)
cat("\nFrom labware (TEXT_ID):\n")
unique(labware_example$TEXT_ID)
cat("\nFrom biota_samples (SAMPLE_ID):\n")
unique(biota_example$SAMPLE_ID)
```


## 8. Add columns from BIOTA_SAMPLES to the Excel data   
I.e., for the the Excel rows where the LIMS number doesn't fit with TEXT_ID in LABWARE_CHECK_SAMPLE, we use
SAMPLE_ID from BIOTA_SAMPLES instead.  

### Make modified LIMS for join  
As in part 3 above ('LIMS_modified'), so we call it 'LIMS_modified2'
```{r}
# In Excel file (LIMS):  "NR-2015-230236" "NR-2015-230559" "NR-2015-230583" "NR-2015-230678" "NR-2015-230611"
# In biota_samples (SAMPLE_ID): 230236 230323 230288 230722 230325 230704 230559 230611 230743 230678 230324

dat <- dat %>% 
  mutate(LIMS_modified2 = case_when(
    Year >= 2015 & nchar(LIMS) == 14 ~ as.numeric(substr(LIMS, 9, 14)),
    TRUE ~ as.numeric(NA)    
  ))

```

### Get BIOTA_SAMPLES data
```{r}
df_samples <- get_nivabase_selection(
  "SAMPLE_ID, SPECIES_ID, TISSUE_ID, STATION_ID, TAXONOMY_CODE_ID, SAMPLE_DATE", 
  "BIOTA_SAMPLES",
  "SAMPLE_ID", 
  dat %>% filter(!is.na(LIMS_modified2)) %>% pull(LIMS_modified2) %>% unique()
  )  # 452 lines

```

```{r}
df_samples %>%
  count(SPECIES_ID, TAXONOMY_CODE_ID, TISSUE_ID)
```

### Add species, tissue and station to 'df_samples'   
Get species, tissue and station using the codes (TAXONOMY_CODE_ID etc.) and join to data
```{r}
df_taxcode <- get_nivabase_selection(
  "TAXONOMY_CODE_ID, NAME", 
  "TAXONOMY_CODES",
  "TAXONOMY_CODE_ID", 
  unique(df_samples$TAXONOMY_CODE_ID)
  )  %>%
  rename(BIOTA_Species = NAME)   # 3 lines

#
# Next lines are not needed, we need only use TAXONOMY_CODE_ID here
#
# df_species <- get_nivabase_selection(
#   "*", 
#   "SPECIES",
#   "SPECIES_ID", 
#   unique(df_samples$SPECIES_ID)
#   )  # 1 line
# 

df_tissue <- get_nivabase_selection(
  "TISSUE_ID, TISSUE_NAME", 
  "BIOTA_TISSUE_TYPES",
  "TISSUE_ID", 
  unique(df_samples$TISSUE_ID)
  ) %>%
  rename(BIOTA_Tissue = TISSUE_NAME)

# For statoins, there are several names (and sometimes codes) per STATION_ID, so they must be collapsed
df_stations <- get_nivabase_selection(
  "STATION_ID, STATION_CODE, STATION_NAME", 
  "PROJECTS_STATIONS",
  "STATION_ID", 
  unique(df_samples$STATION_ID)
  ) %>%
  group_by(STATION_ID) %>%                                                      # Collapsing codes and names
  summarise(BIOTA_Stationcode = paste(unique(STATION_CODE), collapse = "; "),
            BIOTA_Stationname = paste(unique(STATION_NAME), collapse = "; ")
  )

# 
nrow(df_samples)    # 452
df_samples <- df_samples %>%
  left_join(df_taxcode) %>%
  left_join(df_tissue) %>%
  left_join(df_stations) %>%
  rename(BIOTA_Sampledate = SAMPLE_DATE) %>%
  select(SAMPLE_ID, BIOTA_Species, BIOTA_Tissue, BIOTA_Stationcode, BIOTA_Stationname, BIOTA_Sampledate)
nrow(df_samples)    # 452

```

### Join data
```{r}
nrow(dat)  # 59574
dat <- dat %>% 
  left_join(df_samples, by = c("LIMS_modified2" = "SAMPLE_ID"))
nrow(dat)  # 59574
```

## 9. Have a look at final result
### Species  
BIOTA_Species adds no new info
```{r}
dat %>%
  filter(Year >= 2015) %>%
  count(Species, SPECIES, BIOTA_Species)
```

### Tissues  
Both TISSUE and BIOTA_Tissue adds no new info
```{r}
dat %>%
  filter(Year >= 2015) %>%
  count(Organ, TISSUE, BIOTA_Tissue)
```




## 10. Make dat2   
Modified data set for new analyses
Will have new values in 'Species' and 'Organ'

### Make 'Species'  
The old 'Species' is renamed
```{r}
dat2 <- dat %>%
  rename(Species_orig = Species) %>%
  mutate(Species = case_when(
    !is.na(SPECIES) & is.na(Species_orig) ~ SPECIES,
    !is.na(SPECIES) & Species_orig %in% c("Fish","Måke","Bird") ~ SPECIES,
    !is.na(SPECIES) & SPECIES != "Ikke angitt" & Species_orig %in% "Ørret" ~ SPECIES,
    TRUE ~ Species_orig
  )
  )
```

Check  
```{r}
dat2 %>%
  filter(Year >= 2015) %>%
  count(Species_orig, Species, SPECIES, BIOTA_Species)
```


### Remake 'Organ'   
Make Organ_new first (for checking)
```{r}
dat2 <- dat2 %>%
  mutate(Organ_new = case_when(
    is.na(Organ) & (grepl("Blod", TISSUE) | grepl("Blod", BIOTA_Tissue)) ~ "blood",
    is.na(Organ) & (grepl("Egg", TISSUE) | grepl("Egg", BIOTA_Tissue)) ~ "egg",
    is.na(Organ) & (grepl("Lever", TISSUE) | grepl("Lever", BIOTA_Tissue)) ~ "liver",
    is.na(Organ) & (grepl("Muskel", TISSUE) | grepl("Muskel", BIOTA_Tissue)) ~ "filet",
    is.na(Organ) & (grepl("Whole soft body", TISSUE) | grepl("Whole soft body", BIOTA_Tissue)) ~ "whole body",
    is.na(Organ) & grepl("Hel organisme", TISSUE) ~ "whole body",
    is.na(Organ) & !is.na(TISSUE) ~ TISSUE,
    TRUE ~ Organ
  )
  )
```


Check Organ
```{r}
dat2 %>%
  filter(Year >= 2015) %>%
  count(Organ, Organ_new, TISSUE, BIOTA_Tissue)
```

Replace 'Organ'  
```{r}
dat2 <- dat2 %>%
  select(-Organ) %>%
  rename(Organ = Organ_new)
```

### Final shaping up dat2  
Reorder columns
```{r}
dat2 <- dat2 %>%
  select(Description:Species_orig, Species,
         Organ_orig, Organ,
         everything())
```

```{r}
dat2 <- dat2 %>%
  ungroup() %>%
  mutate(Class = case_when(
    Species %in% c("Clupea harengus", "Coregonus albula", "Osmerus eperlanus",
                   "Salmo trutta") ~ "Fish",
    Species %in% c("Euphausiacea indet", "Mysis relicta", "Mytilus edulis",
                   "Pandalus borealis", "Polychaeta") ~ "Aquatic",
    TRUE ~ Class,
  ))
```

## 11. Save
```{r}
# Save in Rdata format
saveRDS(dat2, file = "Data/04_dat2.rds")
# For loading:
# dat2 <- readRDS(file = "Data/04_dat2.rds")

# Save in excel format
infostring <- c(
  "Based on '001 PFAS stacked data table.xlsx'. Column 'Species' and 'Organ' are updated for years",
  "2015 onwards.",
  "'Species' was updated based on column SPECIES (from table 'LABWARE_CHECK_SAMPLE) and from column",
  "'BIOTA_species' (from table 'BIOTA_SAMPLES').",
  "'Organ' was updated based on column TISSUE_NAME (from table 'LABWARE_CHECK_SAMPLE) and from column",
  "'BIOTA_Tissue' (from table 'BIOTA_SAMPLES').",
  "Code: 04_Get_Labware_metadata.Rmd"
)
  
data_for_excel <- list(
  dat2,
  tibble(Info = infostring)
)
names(data_for_excel) <- c("Data", "Info")

openxlsx::write.xlsx(data_for_excel, file = "Data/PFAS stacked data table ver 02.xlsx")

```

## 12. Prepare for making wide format  
Wide format: One column = one parameter  
Needs to have only one row for each combinations of LIMS, PFAS, Species, etc. That is what we do below.  

### Same LIMS can have two different Matrix  
Example below. So we need to group on Matrix as well. 
```{r}
dat2 %>% filter(LIMS %in% "NR-2013-2612-1A" & PFAS %in% "PFBS") %>% select(LIMS, Matrix, PFAS, Data)
```

### Check duplicates 1   
ca 783 duplicates, but as we will see in the next part 
```{r}
#  Check whether we have duplicates
check1 <- dat2 %>%
  count(LIMS, PFAS, Matrix) %>%
  mutate(LIMS_PFAS_Matrix = paste(LIMS, PFAS, Matrix, sep = "_"))

table(check1$n)
```

### Check duplicates 2   
check2: pick duplicates from check1 and list all different values for each variable  
Exploration of the 4 different "cases" in check2  
```{r}
# Function for listing all different variables
vals <- function(x) {
  paste(unique(x), collapse = ";")
}
# vals(c(1,4,4,6,8,19,8))

check2 <- dat2 %>%
  mutate(LIMS_PFAS_Matrix = paste(LIMS, PFAS, Matrix, sep = "_")) %>%
  filter(LIMS_PFAS_Matrix %in% subset(check1, n > 1)$LIMS_PFAS_Matrix) %>%
  select(LIMS, Description, Project, Species, Unit, Count, Year, Matrix, Organ, Class,
         PFAS, Data) %>%
  group_by(LIMS, PFAS) %>%
  summarise_all(vals)

sel <- grepl(";", check2$Data); sum(sel)  # 63, so most of the 773 duplicates have the same value
mean(sel)  # 8 %

# Look at the whole thing (code below is based on this)
# View(check2[sel,])

# Exploration of each "case":

# 1. Some Milkys PFPA data 2006-2007 - two different 'Label_original': PFPA and PFPeA
check2[sel,] %>% filter(PFAS == "PFPA") %>% select(LIMS, PFAS, Description, Species, Class, Data)
dat2 %>% 
  filter(LIMS %in% "NR-2006-60801" & PFAS == "PFPA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)

# 2. Store innsjøer 2015 - registered  twice with slightly different concentrations
#    Sometimes also as species Osnerus eperlanus and Krøkle (same thing)
check2[sel,] %>% 
  filter(Project == "Store innsjøer") %>% 
  select(LIMS, PFAS, Description, Species, Class, Data)
# Used later
lims_storeinnsjo_2015 <- check2[sel,] %>% 
  filter(Project == "Store innsjøer") %>% 
  select(LIMS, PFAS, Description, Species, Class, Data) %>%
  pull(LIMS) %>% unique()
dat2 %>% 
  filter(LIMS %in% "NR-2015-09882" & PFAS == "PFDA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)
dat2 %>% 
  filter(LIMS %in% "NR-2015-09882" & PFAS == "PFDA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)
dat2 %>% 
  filter(LIMS %in% "NR-2015-09889" & PFAS == "PFNA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)

# 3. Tyrifjorden 2018 - some with LIMS number "NR-2018-" (but Description differs)
check2[sel,] %>% 
  filter(Project == "Tyrifjorden") %>% 
  select(LIMS, PFAS, Description, Species, Class, Data)
dat2 %>% 
  filter(LIMS %in% "NR-2018-" & PFAS == "PFDA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)

# 3. Ransfjorden (I964, I965, I969) 2018 - duplicates of PFHpA and Sum
#    Sum: may be with and without detection limit?
#    PFHpA - two different 'Label_original': PFHpA and HPFHpA
check2[sel,] %>% 
  filter(Project == "MilKys" & Description %in% c("I964","I965","I969")) %>% 
  select(LIMS, PFAS, Description, Species, Class, Data)
#    Sum: may be with and without detection limit?
dat2 %>% 
  filter(LIMS %in% "NR-2018-242373") %>% 
  select(LIMS, PFAS, Label_original, Description, Data)
# Used later
lims_ranfjorden <- dat2 %>% 
  filter(Project == "MilKys" & Description %in% c("I964","I965","I969") & PFAS %in% "Sum") %>%
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data) %>%
  pull(LIMS) %>% unique()
#    PFHpA - two different 'Label_original'
dat2 %>% 
  filter(LIMS %in% "NR-2018-242373" & PFAS %in% "PFHpA") %>% 
  select(LIMS, PFAS, Label_original, Description, Species, Class, Data)
check2[sel,] %>% 
  filter(Project == "MilKys" & Description %in% c("I964","I965","I969"))
```

### Duplicates 3: Check PFPA and PFPeA, plus PFHpA and HPFHpA  
PFPeA only found 6 times in 2006-07
HPFHpA only found 9 times in 2018
```{r}
xtabs(~Year + Label_original, subset(dat2, PFAS %in% c("PFPA", "PFPeA")))
xtabs(~Year + Label_original, subset(dat2, PFAS %in% c("PFHpA", "HPFHpA")))
```

### Duplicates 4: Make dat2_corrected
```{r}
# For broad data
dat2_corrected <- dat2 %>%
  filter(!grepl("<", Data)) %>%
  mutate(Data = as.numeric(Data)) %>%
  select(LIMS, Label_original, Description, Project, Species, Unit, Count, Year, Matrix, Organ, Class,
         PFAS, Data)
nrow(dat2_corrected)

# Part 1 and (partly) 4 above - correct PFAS
dat2_corrected <- dat2_corrected %>%
  mutate(PFAS = ifelse(Label_original %in% c("PFPeA", "HPFHpA"), Label_original, PFAS))
         
# Part 4 above: Delete high sums for Ranfjorden blue mussel 
dat2_corrected <- dat2_corrected %>%
  filter(!(LIMS %in% lims_ranfjorden & PFAS %in% "Sum" & Data > 10))
nrow(dat2_corrected)

# Part 2 above - set same Species
# That values differ slighlty we fix by making the means
dat2_corrected <- dat2_corrected %>%
  arrange(LIMS, PFAS, Species) %>%
  group_by(LIMS, PFAS) %>%
  mutate(Species = ifelse(LIMS %in% lims_storeinnsjo_2015, first(Species), Species)) %>%
  ungroup()
nrow(dat2_corrected)

# Part 3 should be no problem, since "Description" differs
```

### Duplicates 5: Make means of the modified data
```{r}
dat2_means <- dat2_corrected %>%
  group_by(LIMS, Label_original, Description, Project, Species, Unit, Count, Year, Matrix, Organ, Class,
           PFAS) %>%
  summarize(Data = mean(Data))
nrow(dat2_means)
```

### Which PFAS has highest percentage detection?
```{r}
# First, make  
detection_rate <- dat2_corrected %>%
  count(PFAS) %>%
  arrange(desc(n))

# Will be used for factor() later
PFAS_ordered <- detection_rate %>% pull(PFAS)

detection_rate
```

## 13. Make wide format data

### All data
```{r}
dat2_wide <- dat2_means %>%
  mutate(PFAS = factor(PFAS, levels = PFAS_ordered)) %>%
  select(LIMS, Label_original, Description, Project, Species, Unit, Count, Year, Matrix, Organ, Class,
         PFAS, Data) %>%
  tidyr::spread(key = PFAS, value = Data)

# Save in Rdata format
saveRDS(dat2_wide, file = "Data/04_dat2_wide.rds")

openxlsx::write.xlsx(dat2_wide, file = "Data/002 PFAS split table ver02.xlsx")

```


### Biota data
```{r}
sum(is.na(dat2_wide$Matrix)) # 0
xtabs(~addNA(Species) + Matrix, dat2_wide)

openxlsx::write.xlsx(dat2_wide %>% filter(Matrix %in% "biota"),
                     file = "Data/003 Subset PFAS biota ver02.xlsx")

```

### Fish data
```{r}
sum(is.na(dat2_wide$Class)) # 2
xtabs(~Species + addNA(Class), subset(dat2_wide, Matrix %in% "biota"))

openxlsx::write.xlsx(dat2_wide %>% filter(Class %in% "Fish"),
                     file = "Data/004 Subset PFAS fish ver02.xlsx")

```

```{r}


df <- dat2 %>%
  filter(!grepl("<", Data)) %>%
  mutate(Data = as.numeric(Data)) %>%
  select(LIMS, Description, Project, Species, Unit, Count, Year, Matrix, Organ, Class,
         PFAS, Data)
df[c(8616,8949),]
```




## OLD STUFF
```{r}
get_nivabase_data("select TABLE_NAME from ALL_TAB_COLUMNS where OWNER = 'NIVADATABASE' and column_name = 'TAXONOMY_CODE_ID'")  
get_nivabase_data("select TABLE_NAME from ALL_TAB_COLUMNS where OWNER = 'NIVADATABASE' and column_name = 'PROJECT_ID'")  
```

```{r}
get_nivabase_data("SELECT * from NIVADATABASE.TAXONOMY where rownum < 4")

```

### In Excel file  
```{r}
?get_nivabase_selection
```

```{r}
# PROSJEKT is never NA in the labware table...
sum(is.na(df_labware$PROSJEKT))
# ... so we use that as a marker for when there was no join from the labware table (LIMS not equal to any TEXT_ID)




sel_newjoin <- dat %>% filter(Matrix %in% "biota" & Year >= 2015 & is.na(PROSJEKT)

dat_sel %>%
  count(Project, Year, Description)


dat_example <- dat_sel %>%
  filter(Project %in% "MilKys" & Year %in% 2015 & Description %in% "13B")
dat_example
```

### Check same year/station in NIVADATABASE.LABWARE_CHECK_SAMPLE
```{r}
labware_example_yr <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where TEXT_ID like 'NR-2015%' and PROSJEKT like '%MILKYS%'")
xtabs(~PROSJEKT, labware_example_yr)
prosjekt <- unique(df$PROSJEKT)
# prosjekt
labware_example <- labware_example_yr %>% filter(grepl("13B", DESCRIPTION, ignore.case = TRUE))
labware_example
```


### Check same station in BIOTA_ tables   
#### First, get BIOTA data from Nivadatabasen
```{r}
df_projects <- get_projects()   # we call it 'df_projects' (the default name used by 'get_stations_from_project')
# Get a list of the stations in the CEMP_Biota project
df_stations <- get_stations_from_project("CEMP_Biota", ignore.case = FALSE)
# Get all specimens collected at these stations (20 seconds or so)
df_specimens <- get_specimens_from_stationdata(df_stations)
df_2015 <- get_biota_chemistry(
  years = 2015, 
  specimendata = df_specimens, 
  stationdata = df_stations,
  report_samples = TRUE)
```

