---
title: "R Notebook"
output: html_notebook
---


## Packages  
```{r}

library(dplyr)
library(niRvana)
library(ggplot2)
library(tidyr)

# install.packages("ODBC")
source("05_Add_MILKYS_data_functions.R")

```

```{r}
# Input your NIVAbasen username/password
set_credentials()

```


## Data from script 05   
```{r}
dat2_means <- readRDS("Data/05_dat2_means.rds")
dat2_wide <- readRDS(file = "Data/05_dat2_wide.rds")

```


## Tables  

### Dry weight  
```{r}

xtabs(~is.na(Drywt), dat2_means)
xtabs(~addNA(Project) + is.na(Drywt), dat2_means)

```

### Fat percentage    
```{r}

xtabs(~is.na(Fatperc), dat2_means)
xtabs(~addNA(Project) + is.na(Fatperc), dat2_means)

xtabs(~addNA(Species) + is.na(Fatperc), dat2_means)

xtabs(~addNA(Year) + is.na(Fatperc), dat2_means %>% filter(Species == "Gadus morhua"))

```


## Name 'dictionary' (from 01)  
```{r}

df_method_sel <- readxl::read_xlsx("Data/01_methods_pfas.xlsx")


```


## Data from Merete (stacked data)
```{r}

dat <- readxl::read_excel("Input_data/001 PFAS stacked data table.xlsx", guess_max = 59000)
# A few errors in the 'PFASlength' variable

tab <- table(dat$LIMS)
length(tab)  # 2809 samples

samplecodes <- names(tab)
  
```


Check year column  
- Should fit with LIMS column (it does)  

```{r}

check <- as.numeric(substr(dat$LIMS,4,7)) == dat$Year
mean(check)  # should be 1

```


### Find samples in LABWARE_CHECK_SAMPLE  

```{r}

labware_vars <- paste(
  "ACCOUNT_NUMBER, CUSTOMER, PROSJEKT, TISSUE, BIOTA_SAMPLENO, X_BULK_BIO, TEXT_ID,", 
  "DESCRIPTION, AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME")

df_check <- get_nivabase_selection(
  labware_vars,
  "LABWARE_CHECK_SAMPLE",
  "TEXT_ID",
  samplecodes, values_are_text = TRUE)

nrow(df_check)

```
### Check length of codes  

```{r}

samplecodes_found <- samplecodes[samplecodes %in% df_check$TEXT_ID]
samplecodes_notfound <- samplecodes[!samplecodes %in% df_check$TEXT_ID]

cat("Number of characters for samplecodes that were found: \n")
table(nchar(samplecodes_found))   # All

cat("\n\n")
cat("Number of characters for samplecodes that were not found: \n")
table(nchar(samplecodes_notfound))

```


### Make modified Lims code    
- the LIMS code should be 13 characters  
- If 12 characters, insert an extra zero  
```{r}

dat <- dat %>%
  mutate(
    LIMS_modified = case_when(
      nchar(LIMS) == 12 ~ paste0(substr(LIMS, 1, 8), "0", substr(LIMS, 9, 12)),
      TRUE ~ LIMS
    )
    
  )

df_labware <- get_nivabase_selection(
  labware_vars,
  "LABWARE_CHECK_SAMPLE",
  "TEXT_ID",
  unique(dat$LIMS_modified), 
  values_are_text = TRUE)

nrow(df_labware)

dat <- dat %>%
  mutate(
    Found_in_labware = LIMS_modified %in% df_labware$TEXT_ID
  )


```

### Check which projects for which we find result   
- None of Milkys samples found  
- All or almost all Elveovervåkning, Milby, Referanseelver, Tyrifjorden samples found   
- Ca 50% to 75% of Store Innsjøer + Urban Fjord samples found       
```{r}

if (FALSE){
  table(df_labware$PROSJEKT) %>% sort(decreasing = TRUE)
  table(dat$Project, dat$Year)
}

table(dat$Project, dat$Found_in_labware)

```

### Check Milkys example  
- Check Ålesund harbour (28A2) 2018   
- TEXT_ID and LIMS numbers are completely different!  
- Can still use DESCRIPTION field  
```{r}

df_example_labware <- get_nivabase_selection(
  labware_vars,
  "LABWARE_CHECK_SAMPLE",
  "AQUAMONITOR_CODE",
  "28A2", 
  values_are_text = TRUE) %>%
  filter(grepl("2018", TEXT_ID)) %>%
  arrange(TEXT_ID)

df_example_excel <- dat %>%
  filter(Description %in% "28A2" & Year == 2018 & Label_original == "PFOS") %>%
  arrange(LIMS)

df_example_labware$TEXT_ID
df_example_excel$LIMS

df_example_excel$Label_original

```

### TODO

- Check Labware table for those Store Innsjøer + Urban Fjord samples that could  be found   
- Check Excel table for those Store Innsjøer + Urban Fjord samples that could NOT be found  
```{r}

# Store Innsjøer + Urban Fjord 

```



### Missing 'species' in the excel file    
Lots of the 2017-18 data lack species - can we get this from Labware?    
- For non-Milkys data - we find Labware data for the vast majority    
    - 13582 samples out of (13582 + 903) samples  
- For Milkys data - we already have species (and could always find anyway)  

```{r}

cat("Missing 'species' by year: \n")
xtabs(~Year + is.na(Species), dat)

cat("\n\n")
cat("Missing 'species' by year: \n")
xtabs(~Found_in_labware + is.na(Species) + (Project == "MilKys"), dat)

```




